---
title: "Week 6 Problem Set"
author: "Jolie Barga  (github link here)"
date: "March 19 2026"
output: pdf_document
---


1. Start a new document, problemset6.Rmd, in your soc10problemsets repository. (You can download a template [here](https://github.com/Soc10Spring2026/soc10problemsets_templates), but be sure to save it in your own soc10problemsets repo). Create a network based on 5 people in our class! (You can use the "class_activities.csv" file from Canvas if helpful or use the code in the template). Using `igraph`, create a network diagram based on who has worked with who in our class activities this quarter (you may want to use `graph_from_data_frame()` if you read in the csv). Then plot the network. (Note: a network with only 5 people, excluding ties to everyone outside of these 5, is sufficient). __Commit__ your changes. 


```{r q1, echo = TRUE}
library(readr)
library(RCurl)
library(magrittr)
library(dplyr)
library(igraph)

# input url from github
url <- getURL("https://raw.githubusercontent.com/tylermcdaniel/css/main/Data/class_activities.csv")

# download class activities data (from github)
class_activities <- read_csv(url)

five_names <- c("Devir, Lindsey", "Leahy, Olivia", "Kang, Christine", 
                "Ong, Alyssa", "Knowles, Genny")

five_people <- class_activities %>%
  filter(name_1 %in% five_names & name_2 %in% five_names)

# network
net <- graph_from_data_frame(five_people, directed = FALSE)

# plot 
plot(net, vertex.size = 30,
     vertex.color = "lavender",
     vertex.frame.color = NA,
     vertex.label.cex = .7,
     edge.curved = .05,
     edge.arrow.size = .3,
     edge.width = .7,
     edge.color = "gray1")
```

2. Choose a centrality measure, and explain how you would interpret this in the context of the class activities data. Who is/are the most central actor(s) according to this measure? __Commit__ your changes. 

```{r q2, echo = TRUE}

degree(net)

```

I chose degree centrality as my measure and to define from class, degree centrality counts how many edges each node has, so the node with the most edges is most central. In the context of class activities from my plotting, degree centrality identifies who has worked with the most classmates which captures the most active collaborator to this specific group of students and assigned to work with the most amount of students in this specific group of 5 students. Using the degree() function, Christine and Alyssa are tied for the highest degree centrality, each with a score of 4, meaning they have worked with the most other students (a total of 4) in the group; therefore the most central actors based on degree centrality.


3. Now download the IPCC co-authorship data (you can use the code included in the template). Using the IPCC co-authorship data, create a network graph of all co-authors who share at least 2 publications together, excluding those with no network ties, similar to what we do in section \@ref(co-authorship-patterns). Graph the data using eigenvector scores, rather than betweenness scores, to plot the vertex sizes. Does this make a difference in our interpretation of the network? __Commit__ your changes. 


Using eigenvector centrality rather than betweenness centrality shifts which authors appear most central because while betweenness identifies authors who bridge different parts of the network, eigenvector centrality rewards authors connected to other well-connected authors meaning an author with fewer total collaborations can still appear large if their collaborators are themselves highly connected. Looking at the network, authors like Nakicenovic, Riahi, and Carraro appear among the most central nodes, so they are deeply embedded in the core cluster rather than simply their role as bridges in the cluster. This shifts our interpretation of the authors toward a tightly knit inner circle of researchers who repeatedly collaborate with each other. This has implications for IPCC authorship because it suggests that influence in this network is concentrated among a small group of highly interconnected scholars, which may reinforce existing geographic and institutional biases in whose perspectives shape global climate policy if only certain regions and beliefs are represented.

```{r q3, echo = TRUE, message = FALSE}

library(readr)
library(RCurl)
library(magrittr)
library(dplyr)
library(igraph)

# input url from github
url <- getURL("https://raw.githubusercontent.com/tylermcdaniel/css/main/Data/IPCC_co_authorship.csv")

# download authorship data (from github)
ipcc_authors <- read_csv(url, skip = 1)

# replace NA's with 0's
ipcc_authors %<>%
  mutate_all(~replace(., is.na(.), 0))

# reorder column names to match rows
ipcc_authors %<>%
  select(order(colnames(.)))

# turn dataframe into matrix
ipcc_authors %<>%
  select(-Author) %>%
  as.matrix()

# replace 1s with 0s
ipcc_authors[ipcc_authors == 1] = 0

# replace NAs with 0s
ipcc_authors[is.na(ipcc_authors)] = 0

# vector of authors who do not collaborate with others
no_colab <- which(apply(ipcc_authors, 1, sum, na.rm = TRUE) == 0)

# remove non-collaborators
ipcc_authors <- ipcc_authors[-no_colab, -no_colab]

# create network
ipcc_net <- graph_from_adjacency_matrix(ipcc_authors,
                                        mode = "undirected")

# eigenvector
V(ipcc_net)$eigenvector <- evcent(ipcc_net)$vector

# plot network with eigenvector scores
plot(ipcc_net,
     vertex.size = V(ipcc_net)$eigenvector/max(V(ipcc_net)$eigenvector) * 20,
     vertex.color = "lavender",
     vertex.frame.color = NA,
     vertex.label.cex = .5,
     edge.curved = .05,
     edge.width = .1,
     edge.color = "gray1")
```

4. Let's say you want to give one of the IPCC authors a memo to disseminate to the others. Who would you choose? Which measure of centrality would help you arrive at your answer? __Commit__ your changes. 


[Identify a reasonable strategy for dissemenation (1 point). Discuss a reasonable centrality measure that aligns with strategy (0.5 points). Correctly identify the node with highest value for said centrality measure (0.5 points).]

```{r q4, echo = TRUE, message = FALSE}

# input url from github
url <- getURL("https://raw.githubusercontent.com/tylermcdaniel/css/main/Data/IPCC_co_authorship.csv")

# download authorship data (from github)
ipcc_authors <- read_csv(url, skip = 1)

# replace NA's with 0's
ipcc_authors %<>%
  mutate_all(~replace(., is.na(.), 0))

```


5. Finally, plot the IPCC co-authorship data, but change the color/shape of the vertices so that these correspond to authors' gender. Describe what you see. What conclusions can you make about authorship and gender? (Note that you will need the IPCC_cv.csv dataset. The code in the template has combined these datasets). __Commit__ your changes. 

[Correctly define a gender variable, and include an accurate label (1 point). Plot the IPCC co-authorship data with modified color/shape based on gender (0.5 points). Accurately describe the plot (0.5 points). Assess conclusions that can (or can't) be drawn (0.5 points).]


```{r q5, echo = TRUE, message = FALSE}

library(readr)
library(RCurl)

# input url from github
url <- getURL("https://raw.githubusercontent.com/tylermcdaniel/css/main/Data/IPCC_co_authorship.csv")

# download authorship data (from github)
ipcc_authors <- read_csv(url, skip = 1)

# replace NA's with 0's
ipcc_authors %<>%
  mutate_all(~replace(., is.na(.), 0))

## now gather cv data
# input url from github
url <- getURL("https://raw.githubusercontent.com/tylermcdaniel/css/main/Data/IPCC_cv.csv")

# download authorship data (from github)
ipcc_cvs <- read_csv(url)


```
