---
title: "Week 5 Problem Set"
author: "Jolie Barga (github link here)"
date: "May 9 2026"
output: pdf_document
---


1. Start a new document, problemset5.Rmd, in your soc10problemsets repository. (You can download a template [here](https://github.com/Soc10Spring2026/soc10problemsets_templates), but be sure to save it in your own soc10problemsets repo). Read the North Carolina shapefile into R using the process described in section \@ref(using-simple-features-in-r). Then plot the data using `ggplot2`. Use a variable other than `AREA` to fill the shapes, and try using one of the `scale_fill_` arguments to create a red color scale. See [these notes on color scales with ggplot](https://ggplot2-book.org/scales-colour.html) for some theory and ideas here. __Commit__ your changes. 



I loaded the North Carolina shapefile using st_read() and plotted it 
using ggplot() and geom_sf(). Instead of using AREA to fill the shapes 
like the class example, I used PERIMETER which shows how the boundary 
length varies across counties in the state. I used scale_fill_gradient() 
with pink and darkred to create a red color scale where darker red 
counties have longer perimeters. I noticed that coastal counties tend 
to have longer perimeters which makes sense given their irregular 
coastline boundaries compared to inland counties.

```{r q1, echo = TRUE}

library(sf)
library(ggplot2)

# north carolina file
nc <- st_read(system.file("shape/nc.shp", package="sf"))

ggplot(nc)+
  geom_sf(aes(fill = PERIMETER))+
  scale_fill_gradient(low = "pink", high = "darkred")+
  labs(title = "North Carolina Counties by Perimeter",
       fill = "Perimeter")+
  theme_minimal()
```

2. Use the `tidycensus` function to download income data for block groups in Santa Clara County. Plot the demographic data using `ggplot()` and comment on what it shows. __Commit__ your changes. 


I used get_acs() to download median household income data for block 
groups in Santa Clara County for 2020. Looking at the map, there is 
a lot of variation in income across the county. The large eastern and 
southeastern portions of the county show very high incomes in the 
$150,000-$250,000 range shown in the lighter purple and pink tones, 
which likely corresponds to the wealthier hillside and tech corridor 
areas. The more densely populated central and western areas around 
San Jose show much more variation, with darker purple and black areas 
indicating lower incomes around $50,000 or below sitting right next 
to lighter areas with much higher incomes. The gray areas represent 
block groups where data was not available. This pattern reflects the 
well known economic inequality in the Bay Area where being close to 
tech hubs and the high cost of housing create really stark differences 
in income even across very short distances.

```{r q2, echo = TRUE, message = FALSE}

library(tidycensus)
library(ggplot2)

census_api_key("ffea6c5d2bcdc3c8fdf9ff3f6aa412b711e54559")

# Santa Clara County
santa_clara <- get_acs(
  state = "CA",
  county = "Santa Clara",
  geography = "block group",
  variables = "B19013_001",
  geometry = TRUE,
  year = 2020)

# plot
ggplot(data = santa_clara, aes(fill = estimate)) +
  geom_sf(color = NA) +
  scale_fill_viridis_c(option = "magma") +
  labs(title = "Median Household Income in Santa Clara County",
       fill = "Income Estimate") +
  theme_minimal()

```

3. Use the `tigris` function to download spatial data for counties in California. Plot the data using `ggplot()`. __Commit__ your changes. 


```{r q3, echo = TRUE, message = FALSE}

library(tigris)
library(ggplot2)

counties <- counties(state = "CA")

# plot counties
ggplot(counties)+
  geom_sf()+
  labs(title = "California Counties") +
  theme_minimal()

```

4. Download the clean AQI monitor datasets from Canvas ("aqi22_clean_ca.csv") or using the code below, and plot these inside the boundaries that you have downloaded from `tigris`. Then, calculate the average PM2.5 levels in each of the geographical units, and plot your geographies again according to these levels (as we did in section \@ref(spatially-joining-data)). Which areas had the best (and worst) air quality in 2022? What are the implications? __Commit__ your changes. 

The first plot shows the PM2.5 monitor locations across California 
overlaid on county boundaries. The second plot shows the average PM2.5 
levels aggregated to the county level. I can see that the Central Valley 
counties like Fresno had the worst air quality in 2022 while coastal 
and northern counties had the best. This matters because the Central 
Valley has a large Latino agricultural and immigrant working population 
that is lower income, meaning the communities with the worst air quality 
are also the most economically vulnerable which reflects broader patterns 
of environmental injustice. Further, medical care is least advanced and 
accessible for people within these Central Valley regions, further 
representing how those who face health impacts from bad air quality also 
will not receive the best medical care and accessibility to receive 
complete and necessary healthcare.


```{r q4, echo = TRUE, message = FALSE}

library(readr)
library(RCurl)
library(magrittr)
library(sf)
library(viridis)
library(ggplot2)
library(dplyr)
library(tigris)

# input url from github
url <- getURL("https://raw.githubusercontent.com/tylermcdaniel/css/main/Data/aqi22_clean_ca.csv")

# download aqi monitor data (from github)
aqi22m <- read_csv(url)

aqi22m %<>%
  st_as_sf(coords = c("X", "Y"))

# set crs
aqi22m %<>%
  st_set_crs(st_crs(counties))

# limit to PM2.5
aqi22m %<>%
  filter(`Pollutant Standard` == "PM25 24-hour 2012" &
           `State Name` == "California")

# cali state boundary
ca <- states() %>%
  filter(NAME == "California")

# plot monitors inside county boundaries
ggplot(aqi22m)+
  geom_sf(data = ca)+
  geom_sf(data = counties)+
  geom_sf(aes(col = `Arithmetic Mean`))+
  scale_color_viridis(option = "magma", direction = -1)+
  theme_minimal()+
  labs(col = "PM2.5 Average")

# spatial join 
counties %<>%
  st_join(aqi22m)

counties %<>%
  group_by(GEOID)

# average AQI for each county
counties %<>%
  summarise(mn_aqi = mean(`Arithmetic Mean`, na.rm = T))

# plot average PM2.5 by county
ggplot(counties)+
  geom_sf(data = ca)+
  geom_sf(aes(fill = mn_aqi))+
  scale_fill_viridis(option = "magma", direction = -1)+
  theme_minimal()+
  labs(fill = "PM2.5 Average",
       title = "Average PM2.5 Levels by County in California 2022")
```


5. Does the county-level air quality map accurately represent how people experience air quality? Is the Modifiable Areal Unit Problem something that we should be concerned with? __Commit__ your changes. 

[Make a defensible argument about the data's representativeness (0.5 points). Make a defensible argument about the MAUP (1 point). ]

