---
title: "Week 7 Problem Set"
author: "Jolie Barga https://github.com/joliebarga/soc10problemsets"
date: "May 19 2026"
output: pdf_document
---


[Note: the scoring guidelines in brackets can be deleted before submitting. Overall presentation/cleanliness of document (e.g. writing and code presented clearly, no unnecessary messages/warnings) 1 point.]

1. Start a new document, problemset7.Rmd, in your soc10problemsets repository. (You can download a template [here](https://github.com/Soc10Spring2026/soc10problemsets_templates), but be sure to save it in your own soc10problemsets repo). Create a map of the `ames` data, but change the cluster specifications to try to get closer to the "true" neighborhoods shown [here](https://019c932f-6501-a8d5-bac5-df7f6ede5a76.share.connect.posit.cloud/predictions-and-machine-learning.html#fig:amesfigure).  Describe the benefits and limitations of your approach. __Commit__ your changes. 

I modified the k-means clustering algorithm to use 28 cluster centers, which matches the number of neighborhoods defined in the ames dataset. By increasing the number of centers from the default of 3 or 6, the algorithm is able to identify geographic distinctions between groups of houses, producing a map that more closely resembles the neighborhood boundaries shown in the class notes. The main benefit of k-means clustering is that it is simple, fast, and does not require any prior knowledge of what the neighborhoods actually look like or where their boundaries are because the algorithm finds these groupings on its own. However, there are several important limitations to this approach. First, k-means assumes that clusters are about equal in size, which may not reflect the true shape or population density of real neighborhoods in Ames. Second, because k-means starts with randomly assigned centroids, the results can vary from run to run. Third, this model only uses longitude and latitude to form clusters, which means it ignores other features that likely define real neighborhood boundaries, such as housing prices, building types, school districts, or historical development patterns which are important considerations as well.

```{r q1, echo = TRUE}
library(tidymodels)
library(sf)
library(magrittr)

# load ames data
data(ames)

# make ames spatial
ames_spatial <- ames %>%
  st_as_sf(coords = c("Longitude", "Latitude"))

# k-means clustering
kclust <- kmeans(ames %>%
                   select(Longitude, Latitude), centers = 28)

# add clusters
ames_spatial %<>% 
  mutate(cluster = as.character(kclust$cluster))

ggplot(ames_spatial) + 
  geom_sf(aes(col = cluster))
```

2. Using the `tigris` package, download roads data for the county of Story, IA. Add roads to your plot. __Commit__ your changes. 



```{r q2, echo = TRUE}



```

[Successfully download and add roads to the plot (1 point).]


3.  Try to train an algorithm to predict housing sales prices. Create a random forest to predict the sales price, using some new or different variables than the model above. Create a model that _better_ predicts sales price, and explain why it is better. __Commit__ your changes. 

[Create a random forest algorithm that predicts housing sales prices (0.5 points). Modify the model inputs (0.5 points) and describe the changes made (0.5 points). Use metrics to demonstrate why the model is better than the one presented in the course site (0.5 points).]

```{r q3, echo = TRUE, message = FALSE}



```

4. Perhaps the model could be better trained using Cross Validation. Set up a 10-fold cross validation model, and predict sales prices again. Compare these results to your prior results. Are they similar or different, and what does this mean?   __Commit__ your changes. 


[Train a similar model using cross validation (1 point). Correctly compare prediction metrics from cross validated and earlier models (1 point). ]

```{r q4, echo = TRUE, message = FALSE}


```


5. If a real estate company used algorithms to determine home values, could this have any societal implications? Walk the reader through why the data could induce biases and how you might try to resolve these potential problems.  __Commit__ your changes. 

[Identify potential biases to algorithmic determinations of home values, and implications of these biases (1 point). Identify at least one potential way to reduce bias, and explain why it may or may not work (1 point).]


```{r q5, echo = TRUE, message = FALSE}


```
