---
title: "Week 4 Problem Set"
author: "Jolie Barga (github link here)"
date: "April 26 2026"
output: pdf_document
---

1. Start a new document, problemset4.Rmd, in your soc10problemsets repository. (You can download a template [here](https://github.com/Soc10Spring2026/soc10problemsets_templates), but be sure to save it in your own soc10problemsets repo). In the chunk below, show how you would gather data on California's ethnoracial demographics for 2020 using `tidycensus`. (Note: in the chunk below, `eval` is set to false, so you won't see any results. That is ok! Just show the code. And remove your personal key). __Commit__ your changes.

To gather California's ethnoracial demographics for 2020, I used the 
tidycensus package which provides a way to access U.S. 
Census data directly in R and I chose get_decennial() over get_acs() 
because the decennial census is conducted every 10 years and provides 
exact population counts rather than estimates with margins of error, 
making it the most accurate source for racial demographic data for 2020.

I selected five racial variables from the 2020 Decennial Census 
demographic file: White alone (P2_005N), Black or African American 
alone (P2_006N), Asian alone (P2_008N), Hispanic or Latino (P2_002N), 
and Native American (P2_007N). I chose these five groups because they 
represent the major ethnoracial groups in California and align with the 
racial categories used in the ca_dem_table data we worked with in 
Problem Set 3, allowing for consistency across assignments. I used 
geography = "county" to get a breakdown by California county, which 
provides more exact demographic information than a single statewide 
estimate.

```{r q1, echo = TRUE, eval = FALSE}

library(tidycensus)
library(tidyverse)

census_api_key("my_key")

# ca ethnoracial demographics for 2020
ca_race <- get_decennial(
  geography = "county",
  variables = c(
    White = "P2_005N",
    Black = "P2_006N",
    Asian = "P2_008N",
    Hispanic = "P2_002N",
    Native_American = "P2_007N"
  ),
  state = "CA",
  year = 2020,
  sumfile = "dhc"
)

```

2. In Bit by Bit, Salganik describes several possibilities and challenges of survey data in the digital age. Do you think any of these have changed since 2017 (when the book was written)? Pick a new type of survey or data collection, and describe its current possibilities and limitations. __Commit__ your changes. 

In Bit by Bit, Salganik discusses the rise of smartphone surveys as a 
new method for reaching hard-to-reach populations and collecting 
real-time data and one example he highlights is thee study of formerly 
incarcerated men, where smartphones were used to administer daily surveys 
tracking well-being and employment searches after prison release which we discussed in class. While 
this was groundbreaking in 2017, digital data collection 
has shifted dramatically due to tech and ai advancements. 

One major change is the rise of data collection through wearable 
technology like Apple Watches, Fitbits, and Oura Rings. Rather than 
asking people questions, researchers can now continuously collect 
biometric data like heart rate, sleep patterns, stress levels, and 
physical activity without any active participation from the respondent 
after initial consent. Stanford researchers are even is currently conducting a study on Oura Ring usage on college students.
One possiblity this has is that since this method is always-on and non-reactive in the same way 
Salganik describes big data, people often forget they are being monitored and 
behave naturally so therefore wearable tech can capture data that people cannot accurately self-
report, like exact sleep duration or stress responses to daily events, 
which traditional surveys rely on memory and self-perception for. This 
opens up entirely new research questions in medical sociology, mental 
health research, and the study of inequality in physical health outcomes. A limitation is wearables are expensive and disproportionately owned by 
wealthier and younger populations, making samples highly non-
representative and therefore, like the Stanford study, require the researchers to provide the wearable technology as its not frquent people own it already. The data is also extremely sensitive since 
information collected from these is deeply personal and raises serious ethical concerns around 
privacy, consent, and data security. 


3. Gather the data from [this Wikipedia page on incarceration in the US](https://en.wikipedia.org/wiki/Incarceration_in_the_United_States) (you can use the code from Section 4 of the course website to gather this). Use `mutate()`, `parse_number()` from the `readr` package, and the magrittr pipe (`%<>%`) to fix the issue with the `Count` variable. Then plot the Count variable over time with `ggplot()`. __Commit__ your changes.

I gathered the incarceration data from Wikipedia using rvest and then 
used mutate() and parse_number() to fix the Count variable, which was 
stored as a character with commas rather than as a numeric value. The 
plot shows that incarceration counts in the US increased dramatically 
from 1940 through the 2000s before beginning to decline slightly in 
recent years.

```{r q3, echo = TRUE, message = FALSE}

library(rvest)
library(dplyr)
library(magrittr)
library(readr)
library(ggplot2)

us_inc <- read_html("https://en.wikipedia.org/wiki/Incarceration_in_the_United_States")

us_inc_table <- html_nodes(us_inc,
                            xpath = '//*[@id="mw-content-text"]/div[2]/div[5]/table')

us_inc_table %<>%
  html_table()

us_inc_table %<>%
  as.data.frame()

us_inc_table %<>%
  mutate(Count = parse_number(Count))

ggplot(us_inc_table, aes(x = Year, y = Count)) +
  geom_point() +
  labs(title = "US Incarceration Count Over Time",
       x = "Year",
       y = "Count") +
  theme_minimal()

```

4. Now model it! Add a linear model of incarceration over time to your plot. (Note: If you are unable to plot question 3 correctly, you can plot `Rate` as the outcome for this question). __Commit__ your changes.  

I added a linear model to my incarceration plot from Q3 using 
geom_smooth() with method = "lm" which fits a line of best fit through 
the data points. I can see from the plot that the linear model shows an 
overall upward trend in incarceration counts over time, however I notice 
it does not perfectly capture what is actually happening in the data. 
The counts rose very sharply through the 1980s and 1990s and then began 
to decline after around 2008, which a straight line cannot fully 
represent. This makes me think a non-linear model might actually fit 
this data better, which I will explore in Q5.

```{r q4, echo = TRUE}

ggplot(us_inc_table, aes(x = Year, y = Count)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "US Incarceration Count Over Time with Linear Model",
       x = "Year",
       y = "Count") +
  theme_minimal()
```


5. Would a linear or non-linear model be more appropriate for explaining and predicting trends in US society and incarceration? Explain the benefits and drawbacks of each. __Commit__ your changes.  

Looking at the incarceration data, I think a non-linear model is more 
appropriat because he count rose sharply from around 264,000 in 1940 to over 
2.3 million in 2008 and then declined to around 1.7 million by 2020. 
That kind of curve with a clear peak cannot be captured by a straight 
line without removing key data points and necesarry info.

A linear model has the benefit of being simple and easy to interpret because 
the slope gives a clear number for how much incarceration changes 
per year, which makes it easy to communicate findings and make basic 
predictions. However the drawback is that it assumes a constant rate of 
change, which does not match this data at all. A linear model would 
predict incarceration keeps rising indefinitely, which we can see from 
the data is not what actually happened after 2008.

A non-linear model like the model we used in class has the benefit 
of following the actual shape of the data much more closely, capturing 
both the rapid rise through the 1990s and the decline after 2008. This 
gives a much more honest and accurate description of the trend. The 
drawback is that non-linear models are harder to interpret since there 
is no single slope value, and they can overfit the data by following 
small bumps and dips too closely, which makes future predictions less 
reliable and harder to determine.




