---
title: "Week 9 Problem Set"
author: "Jolie Barga https://github.com/joliebarga/soc10problemsets"
date: "May 28 2026"
output: pdf_document
---


1. Start a new document, problemset9.Rmd, in your soc10problemsets repository. (You can download a template [here](https://github.com/Soc10Spring2026/soc10problemsets_templates), but be sure to save it in your own soc10problemsets repo). `RedditExtractoR` to download Reddit comments and threads on data centers. Create a topic model from the subreddit "DataCenterDebate" using $k$ = 5, and show a figure with the 10 relative most common words in each topic. __Commit your changes__. 

[Successfully download comments and threads using the Reddit API (0.5 points). Create a topic model with 5 topics (1 point) and plot the relative most common words in each topic (0.5 points).]

```{r q1, echo = TRUE, message=FALSE, warning=FALSE, eval = FALSE}
library(RedditExtractoR)
library(tidytext)
library(dplyr)
library(magrittr)

# find data center subreddits
dc_subreddits <- find_subreddits("data center")

# get urls from the DataCenterDebate subreddit
dc_urls <- find_thread_urls(keywords = "data center",
                            period = "week")

# extract comments from these pages
dc_comments <- get_thread_content(dc_urls$url)

# look at comments
dc_comments$comments

# look at threads
dc_comments$threads
```

```{r q1-load, echo = TRUE, message=FALSE, warning=FALSE, eval = TRUE}
library(readr)
library(tidytext)
library(dplyr)
library(magrittr)
library(topicmodels)
library(ggplot2)

# load pre-downloaded reddit data on data centers
dc_comments <- read_rds("dc_comments.rds")

# create tidytext datasets
tidy_comments <- dc_comments$comments %>%
  unnest_tokens(word, comment) %>%
  anti_join(stop_words)

tidy_threads <- dc_comments$threads %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words)

tidy_threadtitles <- dc_comments$threads %>%
  unnest_tokens(word, title) %>%
  anti_join(stop_words)

# create counts of each word per document
tidy_comments_counts <- tidy_comments %>%
  count(url, word, sort = TRUE)

tidy_threads_counts <- tidy_threads %>%
  count(url, word, sort = TRUE)

tidy_threadtitles_counts <- tidy_threadtitles %>%
  count(url, word, sort = TRUE)

# combine and remove NA values
tidy_counts <- bind_rows(tidy_comments_counts,
                         tidy_threads_counts,
                         tidy_threadtitles_counts)

tidy_counts %<>%
  na.omit()

# recast as document-term matrix
dc_dtm <- tidy_counts %>%
  cast_dtm(url, word, n)

# run lda with k = 5 topics
dc_lda <- LDA(dc_dtm, 5, control = list(seed = 1))

# get word-topic probabilities
dc_topics <- tidy(dc_lda, matrix = "beta")

# get top 10 words per topic
dc_top_terms <- dc_topics %>%
  group_by(topic) %>%
  slice_max(beta, n = 10) %>%
  ungroup() %>%
  arrange(topic, -beta) %>%
  mutate(term = reorder_within(term, beta, topic))

# plot top words in each topic
ggplot(dc_top_terms, aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  scale_y_reordered()
```

2. Describe the topics that you found in question 1. How do these compare to the topics in the $k$ = 2 case (in section \@ref(topic-models)), and how well do these describe your data? __Commit your changes__.  

The five topics from my k = 5 model seem to capture different themes in the DataCenterDebate subreddit. Topic 1 seems focused on energy and infrastructure, with top words like "gas," "energy," "water," "power," and "natural." This suggests people are discussing the resources that data centers consume. Topic 2 appears to be about AI and industry growth, with words like "ai," "power," "people," "tech," "companies," "building," and "build." This looks like conversations about the tech industry expanding data center capacity across the globe and whats needed/the constituients to do so . Topic 3 looks like it is about local environmental impact, with words like "water," "cooling," "impact," "heat," and "city." This points to community level concerns about how data centers affect local environments and the environmental aspects and impacts it has. Topic 4 seems focused on local news and specific projects, with words like "2025," "city," "news," "bessemer," "project," and "cooling." This suggests coverage of specific proposed or ongoing data center developments in particular places and where the discussion peaked and continues to take place .Topic 5 appears to be a more general discussion topic, with broad terms like "data," "center," "centers," "water," "ai," and "communities." This field combines all of the different charts.
Compared to the k = 2 case from the class notes, the 5 topic model is a lot more informative. In the k = 2 model, the two topics were very broad and hard to distinguish from each other since both had words like "data," "centers," and "water" near the top. It was difficult to say what was really different between the two topics. With k = 5, the topics are much more separated and easier to interpret. For example, Topic 4 picks up on very specific local coverage with words like "bessemer" and "news" that would have been buried inside one of the two broad topics in the k = 2 model. Similarly, Topic 2 pulls out AI and industry language more clearly, whereas in the k = 2 model that kind of language was mixed in with more general data center terms. Overall, the k = 5 model does a better job capturing the range of concerns people raise on this subreddit, from energy use to local community impact to industry growth, even if some topics like Topic 5 are still a bit broad and general.


3. Create an account in [Claude](https://claude.com/product/overview) or [ChatGPT](https://chatgpt.com/) (or both). The free versions are sufficient! Ask your LLM to write R code that downloads Reddit data, models topics (any number) and describes these topics. List the prompts that you used, your impressions of the results, and how these differed (or didn't differ) from your responses to questions 1 and 2.  __Commit your changes__. 




4. Considering Engzell and Wilmers' multi-agent workflow, try to improve your Claude modelling process (you might not be able to set up a true multi-agent workflow in the free versions, but you could simulate this by giving your LLM instructions to review or improve the existing output in specific ways, having ChatGPT review the Claude code or vice versa, etc.). Explain what you tried (including additional prompts), how your results changed (or didn't change), and whether this improved your understanding of the data or not. __Commit your changes__. 


```{r q4, echo = TRUE, message = FALSE}


```

[Describe a new workflow involving multiple LLM queries or agents (1 point). Describe how the results change (0.5 points) and the degree to which this changes understanding of the data (0.5 points).]


5. Consider how LLMs (specifically, their ability to write large amount of text and code with a single prompt) will change the social sciences. Will this technology free up researchers to pursue bigger ideas, or will it lead to "mediocrity at scale?" What do you see as the most important opportunities and challenges related to LLMs in the social sciences? (Please refrain from using your LLM to write your answer - I want to know your thoughts!). __Commit your changes__. 

[Make reasonable predictions with references to relevant materials (0.5 points). Describe opportunities and challenges for the future of social science research (1 point).]

